---
title: "Predicting the Water Temperature of Toronto Beaches"
author: Richard Guo
thanks: "Code and data are available at: [https://github.com/Richard-Guo1/predicting-beach-water-temperature](https://github.com/Richard-Guo1/predicting-beach-water-temperature)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
beachData <- readRDS(here::here("data/02-analysis_data/analysis_data.rds"))
```


# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....






# Data {#sec-data}

## Overview

The data used in this paper is derived from Open Data Toronto and is read into this paper through the opendatatoronto library [@opendatatoronto]. The particular data set used to analyze the observations made by city staff on the conditions of all guarded Toronto beaches between the months of May and September [@beaches]. All the data analysis was done through R [@citeR] with the aid of the following packages: tidyverse [@tidyverse], fastDummies[@fastdummies], here [@here], dplyr [@dplyr], tibble [@tibble], janitor [@janitor], ggplot2 [@ggplot], and knitr [@knitr]. 

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



.

```{r}
#| label: fig-water-temp-year
#| fig-cap: Water Temperature Quantiles by Year
#| echo: false

ggplot(beachData, aes(x = year, y = waterTemp, group = year)) +
  geom_boxplot(alpha = 0.8) +
  theme_minimal() + 
  labs(x = "Year", y = "Water Temperature (°C)")
```

Talk more about it.

And also planes (@fig-temps). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-temps
#| fig-cap: Relationship between Air Temperature and Water Temperature
#| echo: false
#| warning: false
#| message: false

beachData |> 
  ggplot(aes(x = airTemp, y = waterTemp)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Air Temperature (°C)",
       y = "Water Temperature (°C)")
```

Talk way more about it. 

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.








# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up



### Model justification



# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false


first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(first_model)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}


\newpage


# References


