---
title: "Predicting the Water Temperature of Toronto Beaches"
author: Richard Guo
thanks: "Code and data are available at: [https://github.com/Richard-Guo1/predicting-beach-water-temperature](https://github.com/Richard-Guo1/predicting-beach-water-temperature)."
date: today
date-format: long
abstract: "This paper analyzes the efficacy of using easily accessible weather information to predict the water temperature at a beach. Data recorded by City of Toroto staff stationed at various beachfronts between mid May and mid September is used. A linear regression model was used with air temperature, recent rainfall, and the location of the beach as the causal predictors."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(knitr)
library(modelsummary)
beachData <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
```


# Introduction

My estimand is the efficacy of using easily accessible weather information (such as from watching a news channel) to predict the temperature of the water at a beach. By analyzing this dataset, I aim to replicate a portion of the decision making process for when people decide to swim at a beach. 

The result of the analysis done shows that there is a strong correlation between local air and water temperatures, with an increase in one degree Celsius in the air predicting an increase of 0.359 in the water.

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data discusses how the dataset was obtained and analyzed. @sec-model discusses the model chosen to analyze the data. @sec-results examines the results of the observations. @sec-discussion contains thoughts on swimming location comparisons, as well as weaknesses and next steps. @sec-appendix is an appendix.


# Data {#sec-data}

## Overview

The data used in this paper is derived from Open Data Toronto and is read into this paper through the opendatatoronto library [@opendatatoronto]. The particular data set used to analyze the observations made by city staff on the conditions of all guarded Toronto beaches between the months of May and September [@beaches]. All the data analysis was done through R [@citeR] with the aid of the following packages: tidyverse [@tidyverse], arrow [@arrow], fastDummies[@fastdummies], here [@here], dplyr [@dplyr], tibble [@tibble], janitor [@janitor], ggplot2 [@ggplot], and knitr [@knitr]. 

## Measurement
	
The measurement of each variable in the beach observation dataset was conducted by city of Toronto staff. Thermometers were used to record the temperature of both the water and the air. Some omitted datatypes such as a water fowl count were estimated due to lack of proper equipment. All observations made were then recorded. 

## Outcome variable

The variable we are interested in is the water temperature. @fig-water-temp-year shows that temperatures observed between mid May and September vary from slightly over 0 to slightly over 30 degrees Celsius. The majority of observations are concentrated between 10 and 20 degrees Celsius. 


```{r}
#| label: fig-water-temp-year
#| fig-cap: Water Temperature Quantiles by Year
#| echo: false

ggplot(beachData, aes(x = year, y = waterTemp, group = year)) +
  geom_boxplot(alpha = 0.8) +
  theme_minimal() + 
  labs(x = "Year", y = "Water Temperature (°C)")
```


## Predictor variables

The chosen predictor variables from the dataset are the beach name, the air temperature, and the presence of rain. The rain amount category may have provided more accurate estimates, but was dropped due to lack of proper documentation, discussed more in @sec-appendix. 

The main predictor of water temperature is the surrounding air temperature. @fig-temps shows a strong positive correlation between the two variables. @fig-beach shows that each beach has a different median temperature. @fig-rain-1 shows that there is a lower median air temperature when it is raining while @fig-rain-2 shows a larger variance but similar median on comparison. This leads to the addition of an interaction predictor between rain and air temperature. 

```{r}
#| label: fig-temps
#| fig-cap: Relationship between Air Temperature and Water Temperature
#| echo: false
#| warning: false
#| message: false

beachData |> 
  ggplot(aes(x = airTemp, y = waterTemp)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Air Temperature (°C)",
       y = "Water Temperature (°C)")
```

```{r}
#| label: fig-beach
#| fig-cap: Water Temperature Quantiles by Beach
#| echo: false
#| warning: false
#| message: false

beachData |> 
  ggplot(aes(x = beachName, y = waterTemp, group = beachName)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Beach",
       y = "Water Temperature (°C)") +
  coord_flip()

```

```{r}
#| label: fig-rain
#| fig-cap: Air and Water Temperature with Rain
#| echo: false
#| warning: false
#| message: false


beachData |> 
  ggplot(aes(x = rain, y = airTemp, group = rain)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Rain",
       y = "Air Temperature (°C)") +
  coord_flip()

beachData |> 
  ggplot(aes(x = rain, y = waterTemp, group = rain)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Rain",
       y = "Water Temperature (°C)") +
  coord_flip()

```

# Model {#sec-model}

The model chosen for analyzing the data was a linear regression model. 

## Model set-up

After the data set was downloaded, it was cleaned for only information relevant to the modelling. The 'beachName' column contained only character strings indicating which beach the observations came from. For the model, it was separated into multiple dummy variables using the fastDummies [@fastdummies] package. The base case chosen by alphabetical order is Bluffer's Beach Park. The cleaned data was then run through the LM function from the dplyr [@dplyr] package. 


# Results {#sec-results}

Our results are summarized in @tbl-modelresults. On a day with an air temperature of 0 degrees Celsius and no rain, the expected water temperature at Bluffer's Beach Park is 7.547. Each one degree increment in air temperature yields an expected increase of 0.359 in the water. Days which are raining predict a water temperature of 5.81 Celsius with air temperature of 0, and a sharper increase of 0.451 in water temperature per one degree increase in air temperature. Holding the air temperature and rain indicator fixed, each of the other beaches expect water temperatures between 0.5 degrees colder to 3 degrees warmer. 

We observe an R Squared value of 0.230, which means that the chosen predictor variables are responsible for 23% of the observed variance. While this value is low, it is sufficient to be used as a non-rigorous predictor of water temperature without proper equipment or being on-site. 

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false


first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of water temperature based on air temperature, rain, and location"
#| warning: false
#| message: false

modelsummary(
  list("Water Temperature" =first_model),
  gof_omit = 'AIC|BIC'
)
```


# Discussion {#sec-discussion}

## Beaches, Outdoor Pools, and Indoor Pools {#sec-first-point}

Lifeguards in the city of Toronto operate in indoor pools, outdoor pools, and the beachfront. Beach lifeguards require an extra training regimen due to the larger area they are expected to cover. While the beach remains accessible throughout the entire year, lifeguards are only stationed during specific months. Similarly, most city-owned outdoor pools are only open during the summer. Disregarding the environment and culture, the main difference between swimming in beach and outdoor pool water is that all pool water has been treated. Pool water is chlorinated, and kept between 21 to 24 degrees Celsius.  

## Weaknesses and next steps


\newpage

\appendix

# Appendix {#sec-appendix}


## Additional data details


\newpage


# References {#sec-ref}


